Reanalysis Step 4: Main Effects Outcome Analysis

### 2025-08-28 reanalysis-4-outcome_analysis.qmd
imp vis

## Define Model Formula (Main Effects Only)

@@

kable(pool_diag, caption = "Pooling diagnostics for main effect model")
```

>>>>

## Define Model Formula

@@

kable(pool_diag, caption = "Pooling diagnostics for main effect model")

# Extract lapse coefficient from each imputation
imputation_estimates <- data.frame(
  imputation = 1:n_imputations,
  log_odds = numeric(n_imputations),
  OR = numeric(n_imputations),
  stringsAsFactors = FALSE
)

# Get the lapse coefficient from each model
for (i in 1:n_imputations) {
  # Extract from model summaries
  lapse_row <- model_summaries[[i]][model_summaries[[i]]$term == "ever_lapse_binary", ]

  if (nrow(lapse_row) > 0) {
    imputation_estimates$log_odds[i] <- lapse_row$estimate
    imputation_estimates$OR[i] <- exp(lapse_row$estimate)
  }
}

# Calculate running average
imputation_estimates$running_avg_OR <- cumsum(imputation_estimates$OR) / seq_along(imputation_estimates$OR)

# Display the convergence table (first 10 and last 10 if more than 20)
if (n_imputations > 20) {
  convergence_display <- rbind(
    head(imputation_estimates, 10),
    data.frame(imputation = "...", log_odds = "...", OR = "...", running_avg_OR = "..."),
    tail(imputation_estimates, 10)
  )
} else {
  convergence_display <- imputation_estimates
}

# Format for display
convergence_display_formatted <- convergence_display
if (is.numeric(convergence_display$OR)) {
  convergence_display_formatted$OR <- round(convergence_display$OR, 4)
  convergence_display_formatted$running_avg_OR <- round(convergence_display$running_avg_OR, 4)
  convergence_display_formatted$log_odds <- round(convergence_display$log_odds, 4)
}

kable(convergence_display_formatted,
      caption = "Lapse effect estimates by imputation with running average",
      col.names = c("Imputation", "Log Odds", "OR", "Running Average OR"))
```

### Pooling Convergence Visualization

```{r convergence-plot}
#| fig-height: 6
#| fig-width: 10

library(ggplot2)

# Create the convergence plot
convergence_plot <- ggplot(imputation_estimates, aes(x = imputation)) +
  # Individual OR estimates as points
  geom_point(aes(y = OR), color = "#2E86AB", size = 2, alpha = 0.6) +

  # Running average as a line
  geom_line(aes(y = running_avg_OR), color = "#D62728", size = 1.2) +

  # Add a horizontal line at the final pooled estimate
  geom_hline(yintercept = tail(imputation_estimates$running_avg_OR, 1), 
             linetype = "dashed", color = "#D62728", alpha = 0.5) +

  # Labels and theme
  labs(
    title = "Convergence of Pooled Estimate Across Imputations",
    subtitle = "Blue points = individual imputation OR, Red line = running average",
    x = "Imputation Number",
    y = "Odds Ratio",
    caption = paste0("Final pooled OR: ", round(tail(imputation_estimates$running_avg_OR, 1), 3))
  ) +

  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 11),
    plot.caption = element_text(size = 9, hjust = 0.5)
  ) +

  # Set reasonable y-axis limits
  coord_cartesian(ylim = c(
    min(imputation_estimates$OR) * 0.95,
    max(imputation_estimates$OR) * 1.05
  ))

# Add annotation showing stabilization
if (n_imputations >= 10) {
  # Calculate when estimate stabilized (when running avg changes < 1%)
  running_changes <- abs(diff(imputation_estimates$running_avg_OR) / 
                         imputation_estimates$running_avg_OR[-1])
  stabilized_at <- which(running_changes < 0.01)[1]

  if (!is.na(stabilized_at) && stabilized_at < n_imputations - 5) {
    convergence_plot <- convergence_plot +
      geom_vline(xintercept = stabilized_at, linetype = "dotted", alpha = 0.5) +
      annotate("text", x = stabilized_at, y = max(imputation_estimates$OR) * 0.98,
               label = paste0("Stabilized (~", stabilized_at, ")"),
               hjust = -0.1, size = 3.5)
  }
}

print(convergence_plot)

# Calculate and display convergence statistics
convergence_stats <- data.frame(
  Metric = c(
    "Mean OR across imputations",
    "SD of OR across imputations", 
    "CV of OR (%)",
    "Range of OR",
    "Final running average"
  ),
  Value = c(
    round(mean(imputation_estimates$OR), 4),
    round(sd(imputation_estimates$OR), 4),
    round(100 * sd(imputation_estimates$OR) / mean(imputation_estimates$OR), 2),
    paste0("[", round(min(imputation_estimates$OR), 4), ", ", 
           round(max(imputation_estimates$OR), 4), "]"),
    round(tail(imputation_estimates$running_avg_OR, 1), 4)
  )
)

kable(convergence_stats, caption = "Convergence statistics for lapse effect")
```
